---
title: "ASSIGNMENT"
author: "......"
date: "2023-08-24"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidymodels)
library(tidyverse)
library(ggplot2)
library(datasets)
```
Question 1.
Create a factor variable called days that contains the days of the week.
```{r}
# data set containing days of the week
week_days <-c("Monday", "Tuesday", "Wednesday",
              "Thursday", "Friday", "Saturday", "Sunday");week_days
# Creating a factor variable
week_days <-factor(week_days);week_days
#To test whether a variable is truly a factor or not, we use the class() function
class(week_days)
```
Question 2.
Create an Identity matrix of dimensions 3 by 3.
```{r}
#Creating a 3 by 3 identity matrix
ident_matrix <- base::diag(3);ident_matrix
```
```{r}
#Create another 3 by 3 matrix whose data is a sequence from 1 to 9
matrix_1 <-base::matrix(data = seq(1,9,1), 
                              nrow = 3, ncol = 3);matrix_1
```
Matrix multiplication of the two matrices
```{r}
# Multiplying the two matrices by each other
Multiplied_matrix <-ident_matrix %*% matrix_1
Multiplied_matrix
```
Confirm whether matrix_1 is equal to Multiplied_matrix
```{r}
# Confirming the equality of the two matrices
equal_matrix <- base::all(matrix_1 == Multiplied_matrix)
equal_matrix
# The output is TRUE, confirming that the second matrix, matrix_1 is equal to the output of the Multiplied matrix, since the first matrix, iden_matrix is identity multiplied with the second matrix, the result will be the same as that of the second matrix.
```
Question 3. 
Using the starwars dataset
Create a scatterplot using the base graphics of characters height (x-axis) against mass (y-axis).
```{r}
# Loading the required package
library(dplyr)
#Loading the data set to be used
data("starwars")
#Ploting the scatter plot, using the base graphics
scatter_plot <- base::plot(x = starwars$height,
           y = starwars$mass,
           main = "Scatter plot of Height against Mass",
           xlab = "Height", ylab = "Mass", col = "black", pch = 16)
```


Question 4.
Create a histogram of birth year of "starwars"
```{r}
#Histogram of birth year
attach(starwars)
Hist_birth_year <- graphics::hist(birth_year, main = "Histogram of birth year")
```


Question 5.
Creating the plots using ggplot2 package
```{r}
library(ggplot2)
#Creating a scatter plot of Mass against Height
scatter_plot_1 <-ggplot(data = starwars, mapping = aes(x = height, y = mass
                                                       )) +
  geom_point(na.rm = TRUE, col = "blue") +
  ggtitle("SCATTER PLOT OF MASS AGAINST HEIGHT") +
  theme_minimal();scatter_plot_1
```
```{r}
#Creating a histogram
Hist_birth_year_1 <-ggplot(data = starwars, mapping = aes(x = birth_year)) +
  geom_histogram(na.rm = TRUE) +
  ggtitle("HISTOGRAM OF BIRTH YEAR") +
  theme_minimal();Hist_birth_year_1
```
Question 6.
Create a function that takes in a character's name from the film star wars, and outputs the films in which the character appears.
```{r}
#Creating a list of the name (characters) and their associated films they appeared in.
name_films <- list(
  "Luke Skywalker" = c("The Empire Strikes Back","A New Hope", "Return of the Jedi", "The Force Awakens", "Revenge of the Sith"),
  "Leia Organa" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Force Awakens"),
  "Darth Vader" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"),
  "Han Solo" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Force Awakens"),
  "Yoda" = c("The Empire Strikes Back", "Return of the Jedi"),
  "Obi-Wan Kenobi" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi"),
  "R2-D2" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Force Awakens"),
  "C-3PO" = c("A New Hope", "The Empire Strikes Back", "Return of the Jedi", "The Force Awakens"),"Roos Tarpals" = c("The Phantom menace", "Padme Amidala" = "The phantom menace","Attack of the clones","Revenge of the Sith"),"Captain Phasma" = c("The force Awakens"), "Sly Moore" = c("Attack of the clones","Revenge of the Sith"), "Raymus Antilles" = c("Revenge of the Sith", "A new Hope")
)
#Creating the function that would output the character names that appeared in a film
Films <- function(character_name) {
  films <- name_films[[character_name]]
  if (is.null(films)) {
    return(paste("Character", character_name, "not found in the film(s)"))
  } else {
    return(films)
  }
}
character_name <- "Raymus Antilles"
films_participated <- Films(character_name);films_participated
```
```{r}
character_name <-"Luke Skywalker"
films_participated <- Films(character_name);films_participated
```

